### 2017-06-21 manuscript/Appendix2.Rmd
updates model description with random year effects

\section{Random Slopes, Random Intercepts Model}
\subsection{Model Description}
Our hierarchical Bayesian model with random intercepts and random slopes allowed us to test for differences among treatments in the relationship between ANPP and growing season precipitation, and allowed us to account for the non-independence of observations through time within a plot.
The random effects structure mirrors the nested structure of the data: plot-level parameters are nested within treatment-level parameters, which are nested within overall parameters.
In what follows, $\textbf{X}$ is the design matrix including a column of 1s for intercepts and a column of precipitation values for each observation, $\textbf{y}$ is a row vector of log(ANPP) values for each observation, *i* references observations, *j* references plots, and *k* references treatment.
The notation *i(j(k))* reads as 'observation *i* associated with plot *j* associated with treatment *k*'.
Recall we have three treatments: drought (~50\% decrease in precipitation), control (ambient precipitation), and irrigation (~50\% increase in precipitation).

@@

\begin{align}
y_{i(j(k))} \sim \text{Normal}\left(\mu_{i(j(k))}, \sigma^2_k \right),
\end{align}

\noindent{}where $\mu_{i(j(k))}$ is the determinstic expectation from the regression model,

@@

\begin{align}
\mu_{i(j(k))} = \textbf{x}_i'\boldsymbol{\beta_{j(k)}}.
\end{align}

Regression parameters (the intercept and slope in the parameter vector $\boldsymbol{\beta}$) were modeled hierarchically,

@@

@@

\subsection{Stan Code}
```{r stan1, echo=T, eval=F}
data {
  int<lower=0> Npreds;        # number of covariates, including intercept
  int<lower=0> Nplots;        # number of plots
  int<lower=0> Ntreats;       # number of treatments
  int<lower=0> Nobs;          # number of observations
  int<lower=0> Nppts;         # Number of precip levels to predict
  vector[Nobs] y;             # vector of observations
  row_vector[Npreds] x[Nobs]; # design matrix
  matrix[Nppts,Npreds] newx;  # design matrix for predictions 
  matrix[Npreds,Npreds] R;	  # priors for covariance matrix
  int plot_id[Nobs];          # vector of plot ids
  int treat_id[Nobs];         # vector of treatment ids
}

parameters {
  vector[Npreds] beta_plot[Nplots]; # a unique vector matrix for each plot
  vector[Npreds] beta_treat[Ntreats]; # a unique vector matrix for each treatment
	vector[Npreds] beta_mu; # overall coefficients
	cov_matrix[Npreds] Sigma; # covariance matrix for treatment-level coefficients
	cov_matrix[Npreds] Sigma_plot[Ntreats]; # unique covariance matrix for plot-level coefs by treatment
	vector<lower=0>[Ntreats] sd_y; # treatment-level observation std. dev.
}

transformed parameters {
  vector[Nobs] yhat; # vector of expected values (predictions)
  for (i in 1:Nobs)
    # regression model for expected values (one for each plot-year)
    yhat[i] = x[i]*beta_plot[plot_id[i]]; 
}

@@

	for(i in 1:Ntreats){
	  # treatment-level coefficients vary normally around overall coefficients
	  beta_treat[i] ~ multi_normal(beta_mu, Sigma);

	  # priors on covariance of effects at plot-level
	  Sigma_plot[i] ~ inv_wishart(Npreds+1, R);
	}

	beta_mu ~ normal(0,1); # priors on overall effects
	Sigma ~ inv_wishart(Npreds+1, R); # priors on covariance of effects at treatment-level
	sd_y ~ weibull(2,1); # priors on observation std. dev. for each treatment

>>>>

\section{Random Slopes, Random Intercepts Model}
\subsection{Model Description}
Our hierarchical Bayesian model with random intercepts and random slopes allowed us to test for differences among treatments in the relationship between ANPP and growing season precipitation, and allowed us to account for the non-independence of observations through time within a plot.
The random effects structure mirrors the nested structure of the data: plot-level parameters are nested within treatment-level parameters, which are nested within overall parameters.
In what follows, $\textbf{X}$ is the design matrix including a column of 1s for intercepts and a column of precipitation values for each observation, $\textbf{y}$ is a row vector of log(ANPP) values for each observation, *i* references observations, *j* references plots, *k* references treatment, and *t* references year.
The notation *i(j(k(t)))* reads as 'observation *i* associated with plot *j* associated with treatment *k* within year *t*'.
Recall we have three treatments: drought (~50\% decrease in precipitation), control (ambient precipitation), and irrigation (~50\% increase in precipitation).

@@

\begin{align}
y_{i(j(k(t)))} \sim \text{Normal}\left(\mu_{i(j(k(t)))}, \sigma^2_k \right),
\end{align}

\noindent{}where $\mu_{i(j(k(t)))}$ is the determinstic expectation from the regression model,

@@

\begin{align}
\mu_{i(j(k(t)))} = \textbf{x}_i'\boldsymbol{\beta_{j(k)} + \gamma_{t}}.
\end{align}

The intercept and slope in the parameter vector $\boldsymbol{\beta}$ were modeled hierarchically,

@@

The random year effects ($\boldsymbol{\gamma}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$),

\begin{align}
\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right).
\end{align}

@@

\subsection{Stan Code}
```{r stan1, echo=T, eval=F}
data {
  int<lower=0> Npreds;        # number of covariates, including intercept
  int<lower=0> Nplots;        # number of plots
  int<lower=0> Ntreats;       # number of treatments
  int<lower=0> Nobs;          # number of observations
  int<lower=0> Nppts;         # Number of precip levels to predict
  int<lower=0> Nyears;        # Number of years
  vector[Nobs] y;             # vector of observations
  row_vector[Npreds] x[Nobs]; # design matrix
  matrix[Nppts,Npreds] newx;  # design matrix for predictions 
  matrix[Npreds,Npreds] R;	  # priors for covariance matrix
  int plot_id[Nobs];          # vector of plot ids
  int treat_id[Nobs];         # vector of treatment ids
  int year_id[Nobs];          # vector of year ids
}

parameters {
  vector[Npreds] beta_plot[Nplots]; # a unique vector matrix for each plot
  vector[Npreds] beta_treat[Ntreats]; # a unique vector matrix for each treatment
	vector[Npreds] beta_mu; # overall coefficients
	vector[Nyears] year_off; # vector of year effects
	cov_matrix[Npreds] Sigma; # covariance matrix for treatment-level coefficients
	cov_matrix[Npreds] Sigma_plot[Ntreats]; # unique covariance matrix for plot-level coefficients for each treatment
	vector<lower=0>[Ntreats] sd_y; # treatment-level observation std. dev.
	vector<lower=0>[Nyears] sigma_year; # year std. dev. hyperprior
}

transformed parameters {
  vector[Nobs] yhat; # vector of expected values (predictions)
  for (i in 1:Nobs)
    # regression model for expected values (one for each plot-year)
    yhat[i] = x[i]*beta_plot[plot_id[i]] + year_off[year_id[i]];
}

@@

	for(i in 1:Ntreats){
	  # treatment-level coefficients vary normally around overall coefficients
	  beta_treat[i] ~ multi_normal(beta_mu, Sigma);
	  Sigma_plot[i] ~ inv_wishart(Npreds+1, R); # priors on covariance of effects at plot-level
	}

	beta_mu ~ normal(0,1); # priors on overall effects
	year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
	Sigma ~ inv_wishart(Npreds+1, R); # priors on covariance of effects at treatment-level
	sd_y ~ weibull(2,1); # priors on observation std. dev. for each treatment
	sigma_year ~ weibull(2,1); # hyperprior on year std. dev.
